Flux2022

### 2022-06-07 ABCDModel.Rmd
Added plots, consolidated analyses and figures into one file

Model Prep (Baseline)
```{r}
# Model prep: split, preprocessing, CV ------------------------------------

# Train / test split ------------------------------------------------------

set.seed(42)
df_split_baseline <- initial_split(
  smri_1,
  prop = 0.80,
  # matching age distributions across train and test set
  strata = "interview_age"
)

@@

# report n training, features, testing
df_train_baseline %>%
  summarise(
    n_training = n(),
    p_features = ncol(df_train_baseline)
  ) %>%
  bind_cols(df_validation_baseline %>%
              summarise(n_testing = n()))

# Pre-processing setup ----------------------------------------------------

# define (what we want to do)
preprocess_recipe_baseline <- df_train_baseline %>%
  # predict scan age by all brain features
  recipe(interview_age ~ .) %>%
  # remove near zero variance predictors
  step_nzv(all_predictors()) %>%
  prep() # where it all gets calculated

preprocess_recipe_baseline

# Apply pre-processing ----------------------------------------------------

# juice() will work with training data, `bake()` to apply this to our test data

# apply on train (gives processed value)
df_train_prep_baseline <- juice(preprocess_recipe_baseline)

# apply on validation
df_validation_prep_baseline <- preprocess_recipe_baseline %>% bake(df_validation_baseline)

# pre process on all of forbow
#frb_prep <- preprocess_recipe_baseline %>% bake(frb_select)

# Cross-validation --------------------------------------------------------

# 10 fold cv repeated 10 times
set.seed(42)
train_cv_baseline <- df_train_prep_baseline %>%
  vfold_cv(
    v = 10, 
    repeats = 10, #10
    strata = interview_age
  )

@@

Train and Fit Model (Baseline)
```{r}
# Extreme gradient boosting
# Parallel processing -----------------------------------------------------

library(doParallel)
all_cores <- parallel::detectCores(logical = TRUE)
# for 8 core 16 thread machine, good performance running more than 
# physical but less than all logical
registerDoParallel(cores = all_cores - 6)

# Model specification -----------------------------------------------------

boost_mod <- boost_tree(
  mode = "regression", 
  trees = 1500, 
  tree_depth = tune(), min_n = tune(), loss_reduction = tune(),
  # randomness
  sample_size = tune(), mtry = tune(), 
  # step size
  learn_rate = tune()
  ) %>%
  set_engine("xgboost", 
             objective = "reg:squarederror")

# Model parameters --------------------------------------------------------

set.seed(42)

# filled 6 dimensional tuning space
xgboost_grid <- grid_latin_hypercube(
  min_n(), tree_depth(), loss_reduction(),
  sample_size = sample_prop(),
  # has unknown, finalize with data to find max
  finalize(mtry(), df_train_prep_baseline),
  learn_rate(),
  size = 500 
)

# Model tuning -----------------------------------------------------------

tictoc::tic()

set.seed(42)
xgb_tuned_results <- tune_grid(
  boost_mod,
  interview_age ~ .,
  resamples = train_cv_baseline,
  grid = xgboost_grid,
  metrics = metric_set(mae, rmse, rsq),
  control = control_grid(verbose = FALSE,
                         save_pred = TRUE)
)

tictoc::toc()
beepr::beep(2)

# Pick best model ---------------------------------------------------------

xgb_tuned_results %>%
  # want to minimize MAE
  show_best("mae")

# select parsimonious params within one SE of best model
best_xgb_params <- xgb_tuned_results %>%
  select_by_one_std_err(metric = "mae", maximize = FALSE, tree_depth)

# Fit and save best model -------------------------------------------------

xgb_final_mod <- boost_mod %>%
  finalize_model(best_xgb_params) %>%
  fit(interview_age ~ .,
      data = df_train_prep)

# save mod
saveRDS(xgb_final_mod, file = here::here(
  "model", "xgboost_abcd_brain_age_mod.rds")
  )

>>>>

Data Import/Set-Up
```{r}
#load("Sample_1.Rds")
#load("Sample_2.Rds")
#load("renamed_smri_baseline.Rda")

#df_train_baseline <- filter(renamed_smri_baseline, src_subject_id %in% sample_1$src_subject_id)
#df_validation_baseline <- filter(renamed_smri_baseline, src_subject_id %in% sample_2$src_subject_id)

#save(df_train_baseline, file="df_train_baseline.Rds")
#save(df_validation_baseline, file="df_validation_baseline.Rds")

load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/analysis_sample_baseline.Rds")
load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/training_sample_baseline.Rds")
```

Test Model
```{r}
abcd_brainage_model <- readRDS("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/xgboost_abcd_brain_age_mod_5_9.rds")
```

Predict Brain Age (baseline)
```{r}
brain_age_df_baseline <-
  abcd_brainage_model %>%
  predict(new_data = analysis_sample_baseline) %>%
  mutate(
    # provide the chronological age at time of scan
    truth = analysis_sample_baseline$interview_age
  ) %>%
  # compute the brain age gap by subtracting chronological age from prediction
  mutate(gap = .pred - truth)

Predict Brain Age (follow-up)
```{r}
brain_age_df_followup <-
  xgb_mod %>%
  predict(new_data = renamed_smri_followup) %>%
  mutate(
    # provide the chronological age at time of scan
    truth = renamed_smri_followup$interview_age
  ) %>%
  # compute the brain age gap by subtracting chronological age from prediction
  mutate(gap = .pred - truth)
```

Evaluate accuracy
```{r}
# compute common performance metrics: mae, rsq, rmse
brain_age_df_baseline %>%
  metrics(truth = truth, estimate = .pred)

brain_age_df_followup %>%
  metrics(truth = truth, estimate = .pred)
```

Optional bias correction
```{r}
# bias prediction method, steps in comments,
# previously determined values hardcoded.

#get validation set
set.seed(42)
df_split_baseline <- initial_split(
  training_sample_baseline, 
  prop = 0.80,
  # matching age distributions across train and test set
  strata = "interview_age"
)

@@

describe(df_train_baseline)
describe(df_validation_baseline)

#think I need to now run model on validation set??
abcd_brain_age_correction <-
  abcd_brainage_model %>%
  predict(new_data = df_validation_baseline) %>%
  mutate(
    # provide the chronological age at time of scan
    truth = df_validation_baseline$interview_age
  ) %>%
  # compute the brain age gap by subtracting chronological age from prediction
  mutate(gap = .pred - truth)

#get MAE for hold-out 
abcd_brain_age_correction %>%
  metrics(truth = truth, estimate = .pred)

abcd_bias_mod_baseline <- lm(.pred ~ truth, data = abcd_brain_age_correction)

# extract intercept and slope
#abcd_bias_mod_baseline$coefficients[["(Intercept)"]]
abcd_bias_intercept_baseline <- 7.917289
#abcd_bias_mod_baseline$coefficients[["truth"]]
abcd_bias_slope_baseline <-  0.2017998

# create bias correct data frame
#Baseline
brain_age_corrected_df_baseline <- brain_age_df_baseline %>%
  mutate(
    # corrected brain age prediction
    corrected_pred =  (.pred - abcd_bias_intercept_baseline) / abcd_bias_slope_baseline
  ) %>%
  # corrected corrected brain age gap
  mutate(corrected_gap = corrected_pred - truth)

#Follow-up
brain_age_corrected_df_followup <- brain_age_df_followup %>%
  mutate(
    # corrected brain age prediction
    corrected_pred =  (.pred - bias_intercept) / bias_slope
  ) %>%
  # corrected corrected brain age gap
  mutate(corrected_gap = corrected_pred - truth)

#mine
brain_age_corrected_df_baseline %>%
  metrics(truth = truth, estimate = .pred)

brain_age_corrected_df_baseline %>%
  metrics(truth = truth, estimate = corrected_pred)

brain_age_corrected_df_followup %>%
  metrics(truth = truth, estimate = corrected_pred)

save(brain_age_df_baseline, file="ABCDBaselineBrainAge.Rda")
save(brain_age_corrected_df_baseline, file="ABCDBaselineBrainAgeCorrected.Rda")

save(brain_age_df_followup, file="ABCDFollowUpBrainAge.Rda")
save(brain_age_corrected_df_followup, file="ABCDFollowUpBrainAgeCorrected.Rda")

@@

### 2022-06-07 Flux2022.Rmd
Added plots, consolidated analyses and figures into one file

#Save samples

save(sample_1, file="Sample1.Rda")
save(sample_2, file = "Sample2.Rda")

@@

Model Figures
```{r}
plot_model(baseline_puberty_y_mean_1, type="pred", terms = "youth_mean", colors = "#67322e", title = "Predicted Gap") +
  theme_sjplot()

plot_model(baseline_puberty_p_mean_1, type="pred", terms = "parent_mean", title = "Predicted Gap") +
  theme_sjplot()

>>>>

#Save samples
save(brainage_pds, file="BrainagePDS.Rda")
save(sample_1, file="Sample1.Rda")
save(sample_2, file = "Sample2.Rda")

@@

Model Figures
```{r}
# Youth-Report Puberty
ggplot(sample_1, aes(x = youth_mean, y = corrected_gap)) +
  stat_smooth(method = "lm", col = "#F08000", fill="#F08000") +
  xlab("Pubertal Development (Youth-Report)") +
  ylab("BrainAGE") +
  ggtitle("Predicted Values") +
  theme_minimal()

# Parent-Report Puberty
ggplot(sample_1, aes(x = parent_mean, y = corrected_gap)) +
    stat_smooth(method = "lm", col = "#F08000", fill="#F08000") +
  xlab("Pubertal Development (Parent-Report)") +
  ylab("BrainAGE") +
  ggtitle("Predicted Values") +
  theme_minimal()

#Cognition
ggplot(sample_1, aes(x = nihtbx_totalcomp_agecorrected, y = corrected_gap)) +
    stat_smooth(method = "lm", col = "#F08000", fill="#F08000") +
  xlab("Age-Corrected Cognition Scores") +
  ylab("BrainAGE") +
  ggtitle("Predicted Values") +
  theme_minimal()

#Old Versions
plot_model(baseline_puberty_p_mean_1, type="pred", terms = "parent_mean", title = "Predicted Gap") +
  theme_sjplot()
